\documentclass{article}

\usepackage{amsmath}
\usepackage{minted}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{blindtext}
\usepackage{amssymb}
\usepackage{scrextend}
\usepackage{multicol}
\usepackage{wrapfig}
\usepackage{tikz}

\usepackage{amsthm}

\usepackage[margin=1in]{geometry}
\pagestyle{empty}

\title{\vspace{-3cm}Homework 3}
\author{Yutong Huang (yxh589)}
\date{}

\begin{document}
    \maketitle
    \section*{Problem 1}
    \begin{proof}
        Proof by contradiction.\\*
        Assume $L_{343}$ is decidable $\implies$ There exists an algotithm that decides $L_{343}$. \\*
        Then assume algorithm $ DecideL_{343} $ correctly decides $ L_{343} $, and contruct the following algorithm:
        \begin{algorithm}
            \DontPrintSemicolon
            \caption{$DecideHalt$}
            \SetKwFunction{MPrime}{M'}
            \SetKwFunction{DL}{$DecideL_{343}$}
            \SetKwProg{sub}{TuringMachine}{}{}
            \SetKwInOut{Input}{Input}

            \Input{$M, w$}
            Encode the following Turing machine:\;
            \sub{\MPrime{x}}{
                Run M on w\;
                \If{x=EECS343, or x=is, or x=fun}{
                    \Return true\;
                }\Else{
                    \Return false\;
                }
            }
            \If{\DL{M'}}{
                \Return true\;
            }\Else{
                \Return false\;
            }
        \end{algorithm}\\*
        Case 1: M halts on input w
        \begin{addmargin}[2em]{0em}
            $\implies$ M' halts and accept any string from $\{EECS343,\ is,\ fun\}$ and no other strings.\\*
            $\implies\ DecideL_{343}$ returns true $\rightarrow\ DecideL_{343}$ accepts.\\*
            $ \implies\ DecideHalt$ returns true $\rightarrow\ DecideHalt$ accepts.   
        \end{addmargin}
        Case 2: M does not halt on input w
        \begin{addmargin}[2em]{0em}
            $\implies$ M' does not halt on any input\\*
            $\implies$ M' does not accept any string from $\{EECS343,\ is,\ fun\}$ and no other strings.\\*
            $\implies\ DecideL_{343}$ returns false $\rightarrow\ DecideL_{343}$ rejects.\\*
            $ \implies\ DecideHalt$ returns false $\rightarrow\ DecideHalt$ rejects.\\*
        \end{addmargin}

        Therefore, the algorithm $DecideHalt$ correctly decides if M halts on input w, which is impossible.
        \begin{proof}
            Assume there is an algorithm $Halt(P,I)$ that correctly decides if an algorithm P halts on input I,
            and construct the following algorithm
            \begin{algorithm}
                \DontPrintSemicolon
                \caption{$Z$}
                \SetKwInOut{Input}{Input}
                \Input{x}
                \If{$Halt(x,x)$}{
                    loop forever\;
                }\Else{
                    \Return\;
                }

            \end{algorithm}\\*
            Consider $Z(Z)$\\*
            Case 1: If Z halts on Z, then Z loops forever $\implies$ contradiction.\\*
            Case 2: If Z does not halt on Z, then Z returns $\implies$ contradiction.\\*
            Therefore, there is no algorithm that correctly decides if a program will halt on given input.
        \end{proof}

        By contradiction, then, $DecideL_{343}$ cannot exist $\implies\ L_{343}$ is undecidable.
    \end{proof}
    \newpage
    \section*{Problem 2}
    $L=\{\langle M,q \rangle|q\ is\ a\ useless\ state\ in\ TM\ M \}$\\* 
    Prove $L$ is undecidable.
    \begin{proof}
        Proof by contradiction, assume $L$ is decidable and a machine $DecideL$ decides it.
        We know that for any Turing machine M, for which $L(M)=\phi$, $q_{accept}$ is a useless state.\\*
        Reduce $L$ to $E_{TM}=\{\langle M \rangle |M\ is\ a\ TM\ and\ L(M)=\phi \}$, and consider the following algorithm:
        \begin{algorithm}
            \caption{$ DecideE_{TM} $}
            \SetKwInOut{Input}{Input}
            \Input{M}
            Run $DecideL(\langle M, q_{accept} \rangle )$ \;
            \If{$DecideL$ accepts}{
                \Return true\;
            }\Else{
                \Return false\;
            }
        \end{algorithm}\\*
        Then we have an algorithm $DecideE_{TM}$ that decides $E_{TM}$, which is impossible according in Theorem 5.2 in the book.\\*

        Therefore, by contradiction, $L$ is undecidable.

    \end{proof}
    \section*{Problem 3}
    Consider $B=\{w|w=0x\ for\ some\ x \in A_{TM}, or\ w=1y\ for\ some y \in \overline{A_{TM}}\}$.\\*\\*
    Then $\overline{B}=\{ w|w=\epsilon,or\ w=0x\ for\ some\ x \in \overline{A_{TM}}, or\ w=1y\ for\ some\ y \in A_{TM}\}$\\*\\*
    Mapping reducibility from $B$ to $\overline{B}$:
    $$ f(w)=\left\{
        \begin{aligned}
        1x &  & (w=0x) \\
        0x &  & (w=1x) \\
        \end{aligned}
        \right. $$
    Then $B\leq_m \overline{B}$.
    \begin{proof}
        $\forall w \in B \Rightarrow f(w) \in \overline{B}$:\\*
        Case 1: $w=0x$, then $f(w)=1x$, and $x\in A_{TM} \implies f(w) \in \overline{B}$\\*
        Case 2: $w=1x$, then $f(w)=0x$, and $x\in \overline{A_{TM}} \implies f(w) \in \overline{B}$\\*
        $\forall f(w) \in \overline{B} \Rightarrow w \in B$:\\*
        Case 1: $f(w)=1x$, then $x \in A_{TM}$ and $w=0x \implies w \in B$\\*
        Case 1: $f(w)=0x$, then $x \in \overline{A_{TM}}$ and $w=1x \implies w \in B$
    \end{proof}
    $B$ is undecidable.
    \begin{proof}
        It's obvious that $A_{TM} \leq_m \overline{B}$, and $A_{TM}$ is unrecognizable, therefore $\overline{B}$ is unrecognizable.\\*
        Since $B\leq_m \overline{B}$, we can also deduce that $B$ is unrecognizable, which entails that $B$ is undecidable.
    \end{proof}
 
\end{document}